\papertitle

### 2017-03-23 article-icmc2017.tex
save

\usepackage{tikz}
\usepackage{listings}
\usepackage{subcaption}
\def\papertitle{Temporal dataflows and their application to generative music}
\def\firstauthor{First Author}
\def\secondauthor{Second Author}
\def\thirdauthor{Third Author}

@@

    \subsection{Related works}
    Citer Kyma
    Céu
    Perl
    Boolean dataflows (http://pop-art.inrialpes.fr/~fradet/PDFs/RR8333.pdf)

%    \subsection{Examples}

    \subsection{Temporal formalism}
    This work implicitely uses the OSSIA formalism to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
    The elements of its syntax are presented in fig.~\ref{fig.iscore-example}.

    \begin{figure}[h]
        \centering
        \includegraphics[width=0.45\textwidth]{images/iscore-example.eps}
        \caption{Part of an i-score scenario, showcasing the temporal syntax used. 
            A full horizontal line means that the time must not be interrupted, 
            while a dashed horizontal line means that the time of this time constraint can be interrupted to proceed 
            to the following parts of the score according to an external event.}
        \label{fig.iscore-example}
    \end{figure}

@@

@@

    We introduce a permanent external environment, that contains a mapping of key-value pairs.
    Keys are specified with OSC-like\cite{osc} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
    The values of the environment may vary independently of the execution of the program: for instance elements of the environment may be changed asynchronously through the network, graphical widgets or physical controls.

@@

    It is up to the implementor of nodes of the graph to choose whether a port can only work: 
    \begin{itemize}
        \item When connected explicitely,
        \item With the environment produced from previous nodes,
        \item With the global environment,
        \item With both of the previous.
    \end{itemize}

    This mechanism allows for new patterns during the creation of a dataflow program.
    For instance, instead of accepting data from a single source, a port could specify an input pattern (see OSC 1.1 paper with XPath) such as: \lstinline|/foo.*/bar/b[a-zA-Z]|.
    % TODO pattern-matched addresses on the output: seems hard ? how to match it with a "pattern-matched" input port ??
	%Inner data: reduces the possibilities of connection: everything has to be there (explicit)

@@

    \subsection{Strict}
    An execution of a sub-graph can only happen when all the nodes of the sub-graph are executing.

@@

    We will get the following set of equations representing the behaviour: 
    \begin{itemize}
        \item During $t_0$: $\emptyset$.
        \item During $t_1$: $f_2 \circ f_1 $ or $f_1 \circ f_2$ depending on the node ordering.
        \item During $t_2$: $\emptyset$. 
    \end{itemize}

@@

	\subsection{Delayed}
	A connection between an output and an input is delayed through bufferisation in a queue.

@@

	\subsection{Parallelized and mixed}
	All the sub-graphs are run with the same input environment. 
	The resulting values put in the environment are undefined, but it is guaranteed that all values returned will have been set, which may trigger event handlers.

@@

	\subsection{Replaced}
	All the nodes are run with the same input environment. 
	The resulting values put in the environment are those of the latest node executing in the following dependent nodes.

@@

    \begin{figure}
    	\begin{subfigure}{0.15\textwidth }

    		\begin{itemize}
    			\item $t_0$: $f_1$.
    			\item $t_1$: $f_2 \circ f_1$.
    			\item $t_2$: $f_2$. 
    		\end{itemize}
    		\caption{Glutton execution in the order of the fig.~\ref{fig.simple}}
    	\end{subfigure}~
    \begin{subfigure}{0.15\textwidth}

    	\begin{itemize}
    		\item $t_0$: $f_2$.
    		\item $t_1$: $f_1 \circ f_2$.
    		\item $t_2$: $f_1$. 
    	\end{itemize}
    	\caption{
    		Instead, if we are in the case of fig.~\ref{fig.simple-reversed}, we get: }
    \end{subfigure}
    \caption{}
    \end{figure}

	%\section{Evaluation algorithm}
    %We present here the evaluation algorithm that operates at each tick.

@@

    Explicit cables: optimization.
\end{document}

>>>>

\setdefaultlanguage{english}
\usepackage{tikz}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{enumitem}\setlist{nosep}
\setitemize{itemsep=3pt}

\usepackage[backend=biber]{biblatex}
\addbibresource{biblio.bib}

\def\papertitle{Temporal dataflows and their application to generative music}
\def\firstauthor{First Author}
\def\secondauthor{Second Author}
\def\thirdauthor{Third Author}

@@

\subsection{Related works}

OpenMusic reactive\cite{bresson:hal-00965747}
Timed sequences\cite{garcia:hal-01484077}
Citer Kyma\cite{scaletti1989kyma}
Céu\cite{sant2015structured}
Pearl\cite{halang2001safe}\footnote{Not to be mistaken with the Perl language commonly used for text processing}

Dataflows and especially synchronous dataflows have seen tremendous usage in the music and signal processing community. 
A list of patterns commonly used when developing dataflow-based music software is presented in~\cite{arumi2006dataflow}.
Formal semantics for this programming paradigm are given in~\cite{benveniste_data-flow_1993}.
Specific implementations of dataflow systems are discussed in both editions of the Handbook of Signal Processing Systems\cite{bhattacharyya_handbook_2013}.

Dynamicity in dataflows is generally exprimed on two independent aspects : dynamicity of the data, and of the topology.
The first relates to the variability on the streams of tokens, while the second is about changes to the structure of the graph. 
Boolean parametric dataflows\cite{bempelis2015boolean} have been proposed to solve dynamicity of topology, by introducing conditionals at the edges.

\subsection{Temporal formalism}
This work implicitely uses the OSSIA formalism~\cite{celerier2015ossia} to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
The elements of its syntax are presented in fig.~\ref{fig.iscore-example}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.35\textwidth]{images/iscore-example.eps}
  \caption{Part of an i-score scenario, showcasing the temporal syntax used. 
    A full horizontal line means that the time must not be interrupted, 
    while a dashed horizontal line means that the time of this time constraint can be interrupted to proceed 
  to the following parts of the score according to an external event.}
  \label{fig.iscore-example}
\end{figure}

@@

\begin{figure}
	\centering
	\begin{tikzpicture}[scale=0.5]
	 \node[draw, circle] (f1) at (0, 0) { $f_1$ };
	 \node[draw, circle] (f2) at (3, 0) { $f_2$ }
	    edge [<-] (f1);
	\end{tikzpicture}
	\caption{Another trace of execution with the same processes}
	\label{fig.simple-reversed}
\end{figure}

@@

This work extends it with a permanent external environment, that contains a mapping of key-value pairs.
Keys are specified with OSC-like\cite{Freed09featuresand} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
The values of the environment may vary independently of the execution of the program: for instance elements of the environment may be changed asynchronously through the network, graphical widgets or physical controls.

@@

Let $\mathrm{ticks}$ be the function that gives the number of ticks in an interval.+
For a given execution following fig.~\ref{fig.simple}, the evolution of such parameters may look like :

\begin{table}
	\centering
	\begin{tabular}{c|ccc}
		Address & $t_0$ & $t_1$ & $t_2$ \\
		\midrule
		Global \lstinline|/a|  & $x$ & $x $ & \\
		Local \lstinline|/a|   & $\emptyset$ & & \\	
	\end{tabular}
	\caption{}
\end{table}

It is up to the implementor of nodes of the graph to choose whether a port can work: 
\begin{itemize}
  \item When connected explicitely,
  \item With the environment produced from previous nodes,
  \item With the global environment,
  \item With both of the previous.
\end{itemize}

This mechanism allows for new behaviours during the creation of a dataflow program.
For instance, instead of accepting data from a single source, a port could specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
% TODO pattern-matched addresses on the output: seems hard ? how to match it with a "pattern-matched" input port ??
%Inner data: reduces the possibilities of connection: everything has to be there (explicit)

@@

\subsection{Strict}
In a given tick, an execution of a sub-graph can only happen when all the nodes of the sub-graph are activated.

@@

We will get the following behaviours during each slice: 
\begin{itemize}
  \item During $t_0$: $\emptyset$.
  \item During $t_1$: $f_2 \circ f_1 $ or $f_1 \circ f_2$ depending on the node ordering.
  \item During $t_2$: $\emptyset$. 
\end{itemize}

@@

\subsection{Delay}
A connection between an output and an input is delayed through bufferisation in a queue.

@@

\subsection{Parallelization and mixing}
All the sub-graphs are run with the same input environment. 
The resulting values put in the environment are undefined, but it is guaranteed that all values returned will have been set, which may trigger event handlers.

@@

\subsection{Replacement}
All the nodes are run with the same input environment. 
The resulting values put in the environment are those of the latest node executing in the following dependent nodes.

@@

\begin{figure}
  \begin{subfigure}{0.20\textwidth }    		    		
    \begin{itemize}
      \item $t_0$: $f_1$.
      \item $t_1$: $f_2 \circ f_1$.
      \item $t_2$: $f_2$. 
    \end{itemize}
    \caption{Glutton execution in the order of the fig.~\ref{fig.simple}}
  \end{subfigure}~
  \begin{subfigure}{0.20\textwidth}

    \begin{itemize}
      \item $t_0$: $f_2$.
      \item $t_1$: $f_1 \circ f_2$.
      \item $t_2$: $f_1$. 
    \end{itemize}
    \caption{
    	Glutton execution in the case of fig.~\ref{fig.simple-reversed}}
  \end{subfigure}
  \caption{}
\end{figure}

\section{Evaluation algorithm}
- Split the tick in steps
- Each step, go to the next elements (follow the cables) and find all the nodes that can be currently called.
%We present here the evaluation algorithm that operates at each tick.

@@

Explicit cables: optimization.

Generalization to not only a single variable of time, but any kind of driving parameters
\section{Acknowledgements}
\printbibliography 
\end{document}
